---
title: "A simple analysis of the cereal dataset"
author: "Carl Schwarz"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document:
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
    reference_docx: SampleRmarkdown-Template-STYLE.docx
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
    fig_caption: true
---

```{r setup, include=FALSE}
# Links for RMarkdown help

# Information about RMarkdown is at: http://rmarkdown.rstudio.com

# Chunk options      https://yihui.name/knitr/options/

# Dealing with word templates and Rmarkdown
#    http://stackoverflow.com/questions/41982700/how-to-properly-number-headings-in-word-from-a-rmarkdown-document
#    http://rmarkdown.rstudio.com/articles_docx.html   
#    http://rmarkdown.rstudio.com/word_document_format.html

# Here we open the libraries needed and source any other code as needed
knitr::opts_chunk$set(echo = TRUE)

library(car)
library(emmeans)
library(ggplot2)
library(gridExtra)
library(knitr)
library(pander)
library(plyr)
library(readxl)
library(reshape2)

source('http://www.stat.sfu.ca/~cschwarz/Stat-650/Notes/MyPrograms/schwarz.functions.r')


# Get the raw data and do editing. This won't appear in the report, but gives you 
# an audit trail of what you did to the raw data before creating the rest of the
# template

cereal <- read.csv('cereal.csv', 
              header=TRUE, as.is=TRUE, strip.white=TRUE)

# Define new variables and factors (for categorical variables). CHeck the structure of the data frame
cereal$shelfF <- factor(cereal$shelf)
```


```{r eval=FALSE, echo=FALSE}
# Refer to the cheat sheets for RMarkdown.

# Common problems include:
# - Don't add extra spaces after a periods before a new line. Otherwise the
#   text will break at this point and not run together.
```

# Introduction
What is the relationship between the calories in a serving of breakfast
cereal and the placement of the cereal on the shelf in the supermarket?


# Material and Methods
A sample of `r nrow(cereal)` cereals were sampled from a local grocery store
and the nutritional information (e.g. number of grams of fat, protein, carbohydrates, etc.)
and the number of calories per serving was extracted. 
The display shelf on which the cereal was stored was also recorded.

A single-factor Completely Randomized Design (CRD) analysis of variance was
used to compare the mean number of calories of cereals from each display
shelf. This was followed by a Tukey multiple comparison procedure
to investigate among which shelves the mean may differ.

All computations were performed using `r R.Version()$version.string`.


# Results
The data was screened for outliers and no unusal points were located.

This table summarizes the calories per serving by shelf number.
```{r Table1.summary, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
# Comput the summary statstics for each shelf
cereal$shelfc <- recode(cereal$shelf,
                      " 1='Low'; 2='Middle'; 3='High' ")
xtabs(~shelfc+shelf, data=cereal)

cereal$shelfcF <- factor(cereal$shelfc, levels=c("Low","Middle","High"), order=TRUE)

report <- plyr::ddply(cereal, c("shelfcF"), plyr::summarize,
                   ncereal = length(calories),
                   mean    = round(mean(calories),1),
                   min     = min(calories),
                   max     = max(calories),
                   sd      = round(sd(calories),1))

# Create the output table
colnames(report) <- c(
       "Shelf",
       "n",
       "Mean\ncalories\nper\nserving",
       "Min\ncalories\nper\nserving",
       "Max\ncalories\nper\nserving",
       "SD\ncalories\nper\nserving"
    )
pandoc.table(report,
             caption="Summary statistics on calories per serving",
             justify='lrrrrr',
             split.cells=c(1,1,1,1,1,1))

```

This figure shows a graphical display of the calories per serving

```{r boxplot,echo=FALSE, fig.height=4, fig.width=6, dpi=300}
plot <- ggplot(data=cereal, aes(x=shelfcF, y=calories))+
   ggtitle("Calories by shelf height")+
   geom_point( position=position_jitter(h=.2, w=.2))+
   geom_boxplot(alpha=0.2, outlier.size=0)+
   ylab("Calories per serving")
plot

```

```{r anova,echo=FALSE }
calories.fit <- lm(calories ~ shelfcF, data=cereal)
calories.pvalue <- anova(calories.fit)$"Pr(>F)"[1]

sugars.fit <- lm(sugars ~ shelfcF, data=cereal)
sugars.pvalue <- anova(sugars.fit)$"Pr(>F)"[1]

```

There was no evidence (p=`r round(calories.pvalue,2)`) of a difference in the mean calories per serving
among the shelf height. But there was evidence of a difference in the mean grams of sugar per
serving among the shelf height (p=`r round(sugars.pvalue,2)`). A bar graph of the mean grams
per serving is shown below. There was no evidence that the mean amount of sugar per serving varied
between the *Low* and *High* shelves, but there was evidence that the mean of both shelves
differed from that on the *Middle* shelf.

```{r cld,echo=FALSE }
sugars.lsmo <- emmeans::emmeans(sugars.fit, ~shelfcF)
sugars.cld <- CLD(sugars.lsmo)
sf.cld.plot.bar(sugars.cld, "shelfcF")+
   xlab("Shelf heigh\nwith Tukey comparison")+ylab("Mean grams of sugar (95% ci)")
```



# Summary
We found no evidence that the mean number of calories varied among shelve heights, but there
was evidence the the mean amount of sugars did.


